### 2016-05-15 paper.tex
Major updates following comments by @SilviaOlabarriaga and @vkorkhov

\documentclass[preprint,3p,twocolumn]{elsarticle}

\usepackage[x11names,dvipsnames,table]{xcolor} %for use in color links
\usepackage[subrefformat=parens,labelformat=parens]{subfig}
\usepackage{svg}
\usepackage{url}
\usepackage{flushend}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pbox}
\usepackage{pdfcomment}

\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\newcommand{\todo}[2]{\pdfmargincomment[color=red,author=#1,open=true]{#2}}
\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\closednote}[4]{} % closed notes are hidden. This should obviously be a parameter of \note rather than a separate command.
\newcommand{\answerednote}[4]{\pdfmargincomment[color=green,author=#1-#3,open=true]{#1: #2 \textCR \textCR #3: #4}}
\newcommand{\closedanswerednote}[6]{}

\renewcommand\thesubfigure{\Roman{subfigure}}

\journal{Future Generation Computer Systems}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnotes;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{Software architectures to integrate workflow engines in science gateways}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author[mcgill,creatis]{Tristan Glatard}
\author[mcgill]{Marc-\'Etienne Rousseau}
\author[creatis]{Sorina Camarasu-Pop}

\author[mcgill]{Reza Adalat}
\author[mcgill]{Natacha Beck}
\author[mcgill]{\\Samir Das}
\author[isi]{Rafael Ferreira da Silva}
\author[mcgill]{Najmeh Khalili-Mahani}
\author[stpetersburg]{Vladimir Korkhov}
\author[criugm]{Pierre-Olivier Quirion}
\author[mcgill]{Pierre Rioux}
\author[amc]{S\'ilvia D. Olabarriaga}
\author[criugm]{Pierre Bellec}
\author[mcgill]{Alan C. Evans}

\address[mcgill]{McGill Centre for Integrative Neuroscience, Montreal Neurological Institute, McGill University, Canada.}
\address[creatis]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France.}
\address[criugm]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montr\'eal, QC, Canada.}
\address[isi]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA.}
\address[amc]{Academic Medical Center of the University of Amsterdam, Department of Clinical Epidemiology, Biostatistics and Bioinformatics, Amsterdam, NL.}
\address[stpetersburg]{St. Petersburg State University, Russia.}

\begin{abstract}
  Workflow engines are critical for the efficient and transparent
  exploitation of distributed infrastructures in the ecosystem of
  tools and services offered by science gateways.  We investigate six
  software architectures commonly used to integrate workflow engines
  into science gateways. In \emph{tight integration}, the workflow
  engine shares software components with the science gateway. In
  \emph{service invocation}, the engine is isolated and invoked
  through a specific software interface. In \emph{task encapsulation},
  the engine is wrapped as a computing task executed on the
  infrastructure. In the \emph{pool model}, the engine is bundled in
  an agent that connects to a central pool to fetch and execute
  workflows. In \emph{nested workflows}, the engine is integrated as a
  child process of another engine. In \emph{workflow conversion}, the
  engine is integrated through workflow language conversion. We
  describe and evaluate these architectures with metrics for
  assessment of integration complexity, robustness, extensibility,
  scalability and support for meta-workflows and fine-grained
  debugging. Tight integration and task encapsulation are the easiest
  to integrate and the most robust. Extensibility is equivalent in
  most architectures. The pool model is the most scalable one and
  meta-workflows are only available in task encapsulation, nested
  workflows and workflow conversion. These results provide insights
  for science gateway architects and developers.
\end{abstract}

\begin{keyword}
Workflow engines \sep science gateways \sep software architectures.
\end{keyword}

\end{frontmatter}

%\maketitle

\section{Introduction}

Workflow engines are critical for the efficient and transparent
exploitation of distributed infrastructures in the ecosystem of tools
and services offered by science gateways. Several software
architectures can be adopted to integrate workflow engines in science
gateways, with important consequences on the development effort
required and resulting system.

This paper describes, provides examples and compares such
architectures, based on system-independent representations of their
main components and interactions. It is informed
\closednote{vladimir}{is this a good word for this context,
  'informed'? maybe 'based on'?}{Tristan}{I guess it is the right
  word, keeping it as is.} by our experience in the development and
sustained operation of the CBRAIN~\cite{SHER-14} and
VIP~\cite{GLAT-13} science gateways for medical image analysis during
the past 7 years, as well as by lessons learned from several science
gateway and workflow projects such as
SHIWA\footnote{\url{http://www.shiwa-workflow.eu}} and
ER-flow\footnote{\url{http://www.erflow.eu}}.  In this section we
provide background information and definitions of workflow engines,
science gateways and infrastructures. In
Section~\ref{sec:architectures}, we describe six architectures
\closednote{vladimir}{why is it exactly six? are these architectures
  the ones that happened to be implemented during various projects, or
  is it a full set of all possibilities (I believe not)? I think, it
  must be stated clearly why the focus is given to these particular
  ones - we are reviewing and evaluating the approaches that have been
  used already in our projects}{Tristan}{Fixed, thanks.} that cover
the main patterns that we encountered in the various systems that we
studied during the last few years. The architectures are described
within a consistent framework that underlines the functional
interactions between their main software components. In
Section~\ref{sec:evaluation}, the different architectures are
evaluated with metrics measuring integration complexity, robustness,
extensibility, scalability and other specific features. These metrics
are specifically designed to measure the ability of the architectures
to address workflow-related issues commonly encountered in science
gateways. \closednote{vladimir}{can we explain why exactly these
  metrics are used but not anything else? I think, we need to justify
  that these ones are especially important for the SG case; it would
  be nice to refer to some related research on software architecture
  metrics, Google gives about 2 million pages on this, where you can
  find metrics like cohesion, depth of scenario and so
  on}{Tristan}{The metrics are specifically designed to cover our
  particular workflow-SG context. Do you have any other
  suggestion?} We finally compare the architectures and highlight
recommendations for science gateway architects and 
developers.

\subsection{Workflow engines}

In the last decade, the e-Science community has developed workflow
systems to help application developers access distributed
infrastructures such as clusters, grids, clouds and web
services. These efforts resulted in tools among which
Askalon~\cite{fahringer2005askalon},
Hyperflow~\cite{balis2016hyperflow}, MOTEUR~\cite{GLAT-08i},
Pegasus~\cite{deelman2005pegasus,Deelman201517},
Swift~\cite{zhao2007swift}, Taverna~\cite{oinn2004taverna},
Triana~\cite{taylor2007triana}, VisTrails~\cite{callahan2006managing},
WS-PGRADE~\cite{Kacsuk2012} and
WS-VLAM~\cite{wsvlam}. \closednote{vladimir}{As a past developer of it, I
  would suggest to refer to WS-VLAM~\cite{wsvlam} as well, despite it
  is not very famous, but it played an important role for Amsterdam
  e-science workflows}{Tristan}{Done, thanks.} Such workflow engines usually describe
applications in a high-level language with specific data and control
flow constructs, parallelization operators, visual edition tools,
links with domain-specific application repositories, provenance
recording and other features. An overview of workflow system
capabilities is available in~\cite{deelman2009workflows}.

At the same time, toolboxes have been emerging in various scientific
domains to facilitate the assembly of software components in
consistent ``pipelines''. In neuroimaging, our primary domain of
interest, tools such as Nipype (Neuroimaging in Python, Pipelines and
Interfaces~\cite{gorgolewski2011nipype}), PSOM (Pipeline System for
Octave and Matlab~\cite{bellec2012pipeline}), 
PMP (Poor Man's Pipeline~\cite{Ad-DabbaghY2006}), RPPL~\cite{1174106}, SPM (Statistical
Parametric Mapping~\cite{ashburner2011spm}) and FSL (FMRIB Software
Library~\cite{Jenkinson2012782}) provide abstractions and functions to
handle the data and computing flow between processes implemented in a
variety of programming languages. Such tools were interfaced to
computing infrastructures, in particular
clusters, to execute tasks at a high throughput. Some of
these tools also support advanced features such as provenance
tracking, or  redundancy detection across
analyses to avoid re-computation. A wide array of workflows
have been implemented using these pipeline systems and are now shared across
neuroimaging groups world-wide, which represents a tremendous
opportunity for science gateways to leverage. Domain-specific engines
nicely complement e-Science systems that are more oriented towards the
exploitation of distributed computing infrastructures, in particular
grids and clouds.

In this paper, a
\emph{workflow engine} (also abbreviated \emph{engine}) is a piece of
software that submits interdependent computing \emph{tasks} to an
\emph{infrastructure} (local server, cluster, grid or cloud) based on
a workflow description (a.k.a \emph{workflow}), using input data
that may consist of files, database entries or simple parameter
values. Although simplistic, this definition covers both e-Science workflow systems and
domain-specific pipeline systems.
Some workflow engines, 
usually from the e-Science community, 
may transfer data across the infrastructure, and others, usually
domain-specific ones, may leave this role to external processes. Workflows
may be expressed in any language, including high-level XML or JSON
dialects such as Scufl or Hyperflow, and low-level scripting languages
such as Bash.

\subsection{Science gateways}

\label{sec:intro_sg}

Science gateways are used to share resources within a community and to
provide increased performance and capacity through facilitated access
to storage and computing power. They are often accessible through a
web interface that helps users manage access rights, data transfers,
task execution, and authentication on multiple computing and storage
locations. Workflow engines are part of this ecosystem as core
components to implement and execute
applications.

Various science gateways have been developed, including frameworks
such as Apache Airavata~\cite{marru2011apache}, the Catania Science
Gateway Framework~\cite{ardizzone2012decide} and
WS-PGRADE/gUSE~\cite{Kacsuk2012}. Numerous science gateways were built
using such frameworks~\cite{kacsuk2014science,ardizzone2012decide} or
as standalone systems~\cite{SHER-14,GLAT-13}. Most of these systems
include one or several workflow engines.

% Software properties: integration effort, extensibility, robustness,
% scalability, specific features

Integration between workflow engines and science gateways varies
across systems. Some science gateways are tailored to a particular
engine, while others are more general and host applications
executed by different types of engines. %Some gateways may not even use any
%workflow engine and focus on the execution of simple tasks.

Extensibility is an important property of the integration. New
workflows are added frequently, different types and versions of
workflow engines may be integrated over time, and different kinds of
infrastructure can be targeted.

Scalability is also a crucial concern for such multi-user,
high-throughput systems. For this purpose, science gateways may
balance the load among different instances of the same engine, start
new engines elastically using auto-scaling techniques such as the ones
reviewed in~\cite{lorido2012auto}, and use advanced task scheduling
policies on the infrastructure to improve performance, fault-tolerance
and fairness among users.

Robustness is highly desirable as well since it is key to a
good user experience. Simple architectures facilitate the
implementation effort towards robust interactions which, in turn, have
a positive impact on characteristics such as gateway predictability,
transparency, reliability, traceability and
reproducibility.

Other specific features may also be available, for instance
data vizualisation and quality control, workflow edition, debugging
instruments, or social tools among users.

\subsection{Infrastructure}

Infrastructure consists of the computing and storage
resources involved in workflow execution, as well as the software
services used to access these resources. Infrastructure can be
composed of computing or file servers, databases, clusters, grids or
clouds. Some workflow engines and science gateways may require specific
characteristics, such as the presence of a shared file system between
the computing nodes, the availability of a global task meta-scheduler,
the presence of a file catalog, etc. In the analysis presented below,
such specific requirements are not discussed. Instead, we consider the
infrastructure as an abstract system that can execute tasks and store
data regardless of the enabling mechanisms.

\section{Architectures}
\label{sec:architectures}

Architectures to integrate workflow engines in science gateways are
presented in Figure~\ref{fig:architectures} using the graphical
notations defined in
Figure~\ref{fig:notations}. Table~\ref{table:system-classification}
summarizes the classification of a few systems by architecture.

%http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=4782949

\begin{figure}
\centering
\def\svgwidth{0.8\columnwidth}
\input{figures/notations.pdf_tex}
\caption{Graphical notations}
\label{fig:notations}
\end{figure}

\begin{figure*}
\centering
\hspace*{0.2\columnwidth}
\subfloat[Tight integration]{
\def\svgwidth{0.6\columnwidth}
\input{figures/tight.pdf_tex}
\label{archi:tight}
}
\hfill
\subfloat[Service invocation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/service.pdf_tex}
\label{archi:service}
\hspace*{0.2\columnwidth}
}\\
\hspace*{0.2\columnwidth}
\subfloat[Task encapsulation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/sub-task.pdf_tex}
\label{archi:sub-task}
}
\hfill
\subfloat[Pool]{
\def\svgwidth{0.6\columnwidth}
\input{figures/agent.pdf_tex}
\label{archi:agent}
\hspace*{0.2\columnwidth}
}\\
\subfloat[Nested workflows. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-1.pdf_tex}
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-2.pdf_tex}
\label{archi:nested}
}\\
\subfloat[Workflow conversion. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.5\columnwidth}
\input{figures/import-1.pdf_tex}
\hfill
\def\svgwidth{0.5\columnwidth}
\input{figures/import-2.pdf_tex}
\label{archi:import}
}
\caption{Architectures}
\label{fig:architectures}
\end{figure*}
\begin{table*}
\centering
\footnotesize
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{Systems} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{ardizzone2012decide},
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, Neuroscience Gateway~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{kruger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012},
            Science gateways in~\cite{kacsuk2014science}$^w$
            } \\
  Task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & \pbox{1.5\columnwidth}{SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$, HubZero with Pegasus (via hierarchical workflows)~\cite{Deelman201517}, Tavaxy~\cite{Abouelhoda2012}.
           }\\
  Conversion & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$, Tavaxy~\cite{Abouelhoda2012}, system in~\cite{delaGarza2016}.
\end{tabular}
\caption{Classification of science gateways based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

\subsection{Interactions}

The interactions involved in the architectures are described below and
labeled (\texttt{a},\texttt{b},\texttt{c}, \ldots) as in Figure~\ref{fig:architectures}.

\begin{enumerate}[leftmargin=0cm,itemindent=0.65cm,label=\texttt{(\alph*)}]

\item Workflow integration: consists in adding a new workflow to the
  system so that users can execute it. It is triggered by an
  administrator of the science gateway and it results in an interface,
  for instance a web form, where users can enter the parameters of the
  workflow to be executed. The interaction has two steps:
  (\texttt{a$_1$}) the programs used in the workflow are installed on
  the infrastructure or prepared for on-the-fly installation, which
  may require specific privileges; \closednote{vladimir}{or bundled in
    packages considering all needed dependencies than can be deployed
    on the fly during the workflow execution}{Tristan}{added, thanks!} (\texttt{a$_2$}) the
  workflow is configured in the science gateway so that it becomes
  available to users. Note that integrating a workflow is not the same
  process as integrating a workflow \emph{engine}.
\item Task control: operations to manage tasks on the infrastructure,
  including: authentication, submission, monitoring, termination,
  deletion, etc. Controlling tasks requires to deal with the
  heterogeneous batch managers and meta-schedulers that might be
  available on the infrastructure. When the infrastructure is a grid
  or a cloud, it may for instance be achieved using libraries that
  implement standards such as SAGA (Simple API for Grid
  Applications~\cite{goodale2006saga}), DRMAA (Distributed Resource
  Management Application API~\cite{troger2012distributed}), or OCCI
  (Open Cloud Computing Interface~\cite{edmonds2012toward}).
\item Data control: operations to manage data on the infrastructure,
  such as: upload, download, deletion, browsing, replication, caching,
  etc. Data movements can be triggered by the user in the science
  gateway (\texttt{$c_1$}), to upload input data or download processed
  data. They can also be performed by the workflow engine
  (\texttt{$c_2$}), to transfer workflow data (inputs, outputs, or
  intermediate results) across the infrastructure so that tasks can
  use it. The infrastructure might offer various data storage backends
  with heterogeneous interfaces. 
  Tools and services such as JSAGA~\cite{reynaud2010uniform} or Data Avenue~\cite{hajnal2014data}
  can be used to homogeneize these interfaces.
\item Workflow control: operations to control workflow execution in an
  engine, including: workflow submission, monitoring, termination,
  etc. Workflow control can be coarse-grained (black box) or
  fine-grained (white box). In a coarse-grained model, the various
  tasks created by a workflow execution are masked and the user only
  has a global view of the workflow execution. In a fine-grained
  model, user is exposed to the workflow topology, i.e. to the outputs
  of the individual tasks, their statuses, dependencies and so on.
\item Sub-task control: operations used by tasks to submit sub-tasks
  on the infrastructure, including: submission, monitoring,
  termination, deletion, etc. Sub-task control is similar to
  interaction \texttt{b}, except that information about the parent
  of a sub-task is usually available and used for
  additional control. For instance, the parent task may wait for all
  its sub-tasks to complete before finishing, and conversely all the
  sub-tasks may be terminated if the parent task is killed.
\item Pool-agent: specific to the pool architecture described in
  Section~\ref{sec:pool}. This is an interaction used when agents
  retrieve work from a central pool. It covers agent registration and
  de-registration to the pool, protocols to send work from the pool to
  the agent, mechanisms to update work status, and so on. A similar
  type of interaction is used in pilot-job
  systems~\cite{turilli2015comprehensive} and other agent-based
  computing models.
\item Workflow conversion: translation from one workflow language into
  another. This interaction may not be available or possible for every
  workflow language. It has been developed mostly for translation
  between well-structured and relatively simple workflow languages
  such as GWorkflowDL and Scufl~\cite{OLAB-09}, and for translation
  among the 5 workflow systems in the SHIWA project: Askalon, MOTEUR,
  Taverna, Triana and WS-PGRADE.  
\end{enumerate}

\subsection{Tight integration}

See Figure~\subref*{archi:tight}. The workflow engine is tightly
integrated with the science gateway, which means that it is deployed
on the same machine and potentially shares code, libraries and other
software components with the science gateway. For instance, the
workflow engine might be a portlet in a Liferay portal or a model
in a Ruby on Rails application. The workflow engine and the science
gateway usually share a database where applications, users and other
resources are stored.  In this model, task and
data control are both initiated from the science gateway. Interactions
\texttt{b} and \texttt{c$_2$} are initiated from the workflow engine
while \texttt{c$_1$} comes from other parts of the science gateway,
for instance a data management interface. As in any other model, the
installation of new workflows in the science gateway (\texttt{a$_2$})
and infrastructure (\texttt{a$_1$}) is done by an administrator, for
security reasons. This is the model adopted in the Catania Science
Gateway Framework~\cite{ardizzone2012decide} (see specific documentation on
workflows\footnote{\url{http://bit.ly/1oQrzvQ}}), in the Distributed
application runtime environment
(DARE)~\cite{maddineni2012distributed}, in
Galaxy~\cite{goecks2010galaxy}, and in the LONI Pipeline
Environment~\cite{dinov2009efficient}. Note that tightly integrated
architectures may provide advanced workflow edition features which are
not covered by our analysis.
%The Catania Science Gateway Framework would be a good illustration here (workflow engine is a Java portlet), but I can't find a suitable figure about it.
%CIPRES~\cite{miller2010creating} (on which the Neuroscience gateway
%described in~\cite{sivagnanam2015early} is based) % \note{Tristan}{There is no workflow engine in there.}

\subsection{Service invocation}

\begin{figure*}
\centering
\def\svgwidth{1.5\columnwidth}
\input{figures/VIP.pdf_tex}
\caption{Architecture of the Virtual Imaging Platform (service
  invocation).  Step \texttt{3} corresponds to interaction \texttt{d}
  in Figure~\ref{fig:architectures}(II), step \texttt{2} corresponds to
  interaction \texttt{c$_1$}, step \texttt{4} maps to interaction
  \texttt{b}, and steps 7 and 9 map to \texttt{c$_2$} (in VIP, workflow data transfers are
  performed by the tasks and embedded in their descriptions).}
\label{fig:vip-architecture}
\end{figure*}

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, as a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although in some
variations of the architecture (not represented in
Figure~\ref{fig:architectures}) it might as well be delivered by the
workflow engine directly to the user. \closednote{Sil}{about c1: should it
  also show from engine to infra in the figure? this is what i
  understand from the above sentence}{Tristan}{Actually c1 is either
  coming from the science gatewray or from the workflow engine but
  then it merges with c2. I don't think we can cover both cases in a
  single architecture, the latter case would be another variation of
  this architecture. I modified a bit the text to reflect this.}

This architecture is largely adopted in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the systems
in~\cite{wu2010accelerating,shahand:2012jgc} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture, where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g., pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

\subsection{Task encapsulation}

See Figure~\subref*{archi:sub-task}. The workflow engine is wrapped in
a particular task that can submit sub-tasks to the science gateway
through interaction \texttt{e}. The workflow engine keeps track of the
dependencies between the sub-tasks, but their execution is delegated
to the science gateway that executes them on the infrastructure
through interaction \texttt{b}. Although the science gateway has no
global vision of the workflow, it can keep track of the sub-tasks
submitted by a given task, for instance to cancel them when
the task is canceled. The science gateway may also implement
mechanisms to facilitate the handling of task dependencies, for
instance basic dependency lists as available in most cluster managers
(see for instance attribute \texttt{depend} of option \texttt{-W} in
Torque\footnote{\url{http://docs.adaptivecomputing.com}}).

The science gateway also transfers both user and workflow data across
the infrastructure, so that interactions \texttt{c$_1$} and
\texttt{c$_2$} are both covered by \texttt{c}.

Task encapsulation is implemented in CBRAIN~\cite{SHER-14} where it is
used to integrate the FSL toolkit~\cite{Jenkinson2012782} and the PSOM
workflow engine~\cite{bellec2012pipeline}. The CBRAIN-FSL integration
allows leveraging FSL pipelines written in low-level workflow
languages (Linux executables and scripts) that submit tasks uniformly
through a specific tool called fsl\_sub. This integration is shown in
Figure~\subref*{fig:cbrain-fsl-architecture}, where the science gateway is
represented by components \texttt{CBRAIN portal} and \texttt{CBRAIN
  execution server} and the infrastructure consists of \texttt{Storage
  servers} and \texttt{Computing cluster with shared file system}.

The CBRAIN-PSOM integration~\cite{GLAT-16} is shown in
Figure~\subref*{fig:cbrain-psom-architecture}. The PSOM workflow engine
adopts a pilot-job architecture~\cite{turilli2015comprehensive} where
a master coordinates workflow execution by submitting workers and
establishing direct communication channels with them. Note how this
peculiar execution model is well supported by task encapsulation.

\begin{figure*}
\centering
\subfloat[CBRAIN-FSL integration.]{
\def\svgwidth{\columnwidth}
\input{figures/CBRAIN-FSL.pdf_tex}
\label{fig:cbrain-fsl-architecture}
} \hfill \subfloat[CBRAIN-PSOM integration. Figure
reproduced from~\cite{GLAT-16}.]{\def\svgwidth{\columnwidth}
  \input{figures/CBRAIN-PSOM.pdf_tex
}
\label{fig:cbrain-psom-architecture}
}
\caption{CBRAIN architecture for workflow engine integration
  (task encapsulation).  1: User sends data and workflow execution request to
  storage server(s) and CBRAIN portal. 2: CBRAIN portal sends
  execution request to execution server on cluster. 3: Execution
  server transfers data from storage server(s). 4-5: Execution server
  starts workflow engine (FSL tool or PSOM master) via task
  scheduler. 6: Workflow engine submits sub-tasks to execution server
  (FSL tasks or PSOM agents). 7-8: Execution server starts sub-tasks
  through task scheduler (FSL sub-tasks or PSOM workers). FSL sub-tasks will run locally
  instead of being submitted again to CBRAIN through interaction 6. 8':
  (PSOM only) PSOM master and workers execute workflow. 9. Execution
  server transfers results to storage server(s). Interaction
  \texttt{b} in Figure~\ref{fig:architectures}(III) is implemented by steps
  \texttt{4}, \texttt{5}, \texttt{7} and \texttt{8} (regular
  interactions with batch manager). Interaction \texttt{c} is
  implemented by steps \texttt{3} and \texttt{9}. Interaction
  \texttt{e} is implemented by step \texttt{6} (for FSL: through a
  modified version of the fsl\_sub script available in
  \url{https://github.com/aces/cbrain-plugins-neuro}; for PSOM:
  through a specific development in PSOM).}
\label{fig:cbrain-sub-tasking}
\end{figure*}

\subsection{Pool model}
\label{sec:pool}
See Figure~\subref*{archi:agent}. Workflows are submitted by the
science gateway to a central pool through interaction
\texttt{d}. Agents connect to the pool asynchronously to retrieve and
execute workflows through interaction \texttt{f}. Agents may be
started according to various policies, for instance to ensure load
balancing. Workflow engine controls tasks and data on the
infrastructure through interactions \texttt{b} and \texttt{c$_2$},
science gateway transfers user data through interaction
\texttt{c$_1$}, and administrator installs workflows through
interaction \texttt{a$_1$} and \texttt{a$_2$}.

The pool model was implemented in the SHIWA pool~\cite{ROGE-13}
diagrammed in Figure~\ref{fig:shiwa-pool-architecture}, where agents
can wrap different types of workflow engines to execute workflows
expressed with different languages.

\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/pool-interactions.pdf}
\caption{Architecture of the SHIWA pool. Circle-terminated arrows
  indicate messages that are broadcast to all pool
  clients. Interaction \texttt{d} of Figure~\ref{fig:architectures}(IV) is
  implemented by 3 different calls to the pool: workflow
  submission (\texttt{1} \& \texttt{2}), workflow status retrieval
  (\texttt{0} \& \texttt{3}), and workflow
  retrieval (\texttt{13} \& \texttt{14}). Interaction
  \texttt{f} is implemented through 2 types of calls: workflow
  instance retrieval (\texttt{4}, \texttt{5} and \texttt{6}), and
  workflow instance update (\texttt{11} and \texttt{12}). Workflow
  instance retrieval is used by the agents to fetch work from the
  pool. Workflow instance update is used by the agents to update
  workflow statuses.  Calls \texttt{0}, \texttt{7}, \texttt{8} and
  \texttt{9} are used by
  workflow engine plugins to declare their supported language and to
  launch engines, and by workflow engines to report their status to
  the agent. These calls are specific to the SHIWA Pool implementation
  of the \texttt{agent} component and therefore have no corresponding
  representation in Figure~\ref{fig:architectures}(IV). Figure reproduced
  from~\cite{ROGE-13}.}
\label{fig:shiwa-pool-architecture}
\end{figure*}

\subsection{Nested workflows}

See Figure~\subref*{archi:nested}. In nested workflows
(Figure~\subref*{archi:nested}-Left), a workflow engine is integrated
as a \emph{child} process of a \emph{parent} workflow engine. Parent
and child engines might use different languages and might run on
different infrastructures. A parent workflow is also a
meta-workflow. The science gateway communicates with the parent engine
through abstract interaction \texttt{*$_1$}. The science gateway also
communicates with the infrastructure to transfer user data through
abstract interactions \texttt{*$_2$} and \texttt{*$_3$}. Both workflow
engines communicate with the infrastructure through abstract
interactions \texttt{*$_4$} and \texttt{*$_5$}. The parent engine
communicates with the child engine through abstract interaction
\texttt{*$_6$}. Administrator installs workflows through interactions
\texttt{a$_1$} and \texttt{a$_2$}.

Nested workflows are abstract architectural patterns that can be
instantiated in the various architectures described previously. We
focus on instantiation with the service invocation model
(Figure~\subref*{archi:nested}-Right) as this is the most used
architecture. In the instantiation, we assume that the parent and
child workflow engines are distinct pieces of software that require
different workflow services invoked by distinct \texttt{d}
interactions. If this is not the case, then workflow services can be
collapsed into a single one with a \texttt{d} interaction with
itself. An example of such interaction is the use of hierarchical
workflows in Pegasus~\cite{Deelman201517}.
Workflow engines communicate with infrastructures using
\texttt{b} and \texttt{c$_2$}. Science gateway transfers user data to
infrastructures using \texttt{c$_1$} interactions.

Nested workflows have long been available in workflow engines, for
instance in the Taverna workbench~\cite{oinn2004taverna}. They are
also used implicitly in several platforms where workflow engines are
wrapped in workflow tasks as any other command-line tool. Nested
workflows were notably used by the SHIWA Science Gateway to implement
so-called Coarse-Grained workflow
interoperability~\cite{terstyanszky2014enabling}, i.e. to integrate
various workflow engines in a consistent
platform. Figure~\ref{fig:shiwa-architecture} shows the architecture
used in the SHIWA Simulation Platform for nested workflow execution
with service invocation.
\begin{figure*}
\centering
\includegraphics[width=1.5\columnwidth]{figures/shiwa-science-gateway.pdf}
\caption{Nested workflow execution through the SHIWA Science Gateway. The
  parent workflow engine is \texttt{WS-PGRADE/gUSE}, invoked as a service
  in the Science Gateway (step \texttt{1}, interaction \texttt{d} on
  Figure~\ref{fig:architectures}(V)-Right). Ten different child engines can be used by nested
  workflows, invoked through the \texttt{Submission service} (step
  \texttt{2}, interaction \texttt{d}). Each of these engines can
  submit tasks and transfer data to a distributed computing
  infrastructure (\texttt{DCI}, step \texttt{3}, interactions
  \texttt{b} and \texttt{c$_2$}). User data interactions (\texttt{c$_1$}) and
  application porting ones (\texttt{a}) are not represented. Figure
  reproduced from~\cite{terstyanszky2014enabling} with permission of
  the first author.}
\label{fig:shiwa-architecture}
\end{figure*}

\subsection{Workflow conversion}

See Figure~\subref*{archi:import}. This is an abstract model
instantiated with the service invocation architecture for
consistency. Workflow engines are integrated in the science gateway
through workflow format conversion from a native format to the science
gateway format. \closednote{vladimir}{Are we talking only about importing
  workflows from other wf languages? I would suggest to modify the
  title of the section to 'Workflow conversion and import' then, as
  some gateways use the term 'import' to import their own native
  workflows, like WS-PGRADE, and using import only in the context of
  converting workflows can be confusing}{Tristan}{Good catch, the term
  'import' is actually a bit vague. I guess 'conversion' is more
  accurate. Replaced it everywhere in the text.} Workflow conversion
is usually an offline process that is not involved in workflow
execution. A few systems have implemented workflow conversion. In the
SHIWA Simulation Platform, it is implemented through the IWIR
language, which provides a common language for portability across grid
workflow systems~\cite{plankensteiner-prodan-etal:2013} and allows
conversion among $n$ workflow languages using $2n$ interactions
instead of $n^2$.  Tavaxy~\cite{Abouelhoda2012} enables the import,
merging and execution of Taverna~\cite{oinn2004taverna} and
Galaxy~\cite{goecks2010galaxy} workflows; when some workflow parts
cannot be imported, workflows are run by Tavaxy as nested workflows
using their native engine. The work in~\cite{delaGarza2016} describes
workflow conversion from KNIME~\cite{Berthold2008} to WS-PGRADE/gUSE
and from Galaxy~\cite{goecks2010galaxy} to WS-PGRADE/gUSE.

%Furthermore, both systems provide web-portals allowing users to share and publish their workflows: These are the myExperiment portal for Taverna [23, 24] and the Public Pages for Galaxy [13].

\section{Evaluation}

\label{sec:evaluation}

The architectures described in Section~\ref{sec:architectures} are
evaluated in Table~\ref{table:evaluation} using five main criteria:
integration complexity, robustness, extensibility, scalability and
other specific features. These are based on science gateway properties
introduced in Section~\ref{sec:intro_sg}. Criteria break down to
specific metrics where \emph{lower value indicates better
  performance}. Colors in Table~\ref{table:evaluation} represent
normalized scores, as explained in the Table caption. For each
criterion, a global score is computed by summing up the individual
scores for each metric. We ensure that the different metrics used to
calculate the global score for a criterion are comparable, so that
they can be combined meaningfully. The criteria and metrics are
explained hereafter.

\begin{table*}
\footnotesize
\centering
\begin{tabular}{rcccccc}
                                     & \textbf{Tight}
                                     & \textbf{Service}
                                     & \textbf{Task}
                                     & \textbf{Pool}
                                     & \textbf{Nested}
                                     & \textbf{Conversion} \\
\cellcolor[HTML]{EEEEEE}\textbf{Integration complexity}& \multicolumn{6}{l}{\cellcolor[HTML]{EEEEEE}}\\
  Total components -- \texttt{I$_1$} & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99DD99}3
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99BB99}4
                                     & \cellcolor[HTML]{999999}5
                                     & \cellcolor[HTML]{99DD99}3\\
Total interactions -- \texttt{I$_2$} & \cellcolor[HTML]{99FF99}5
                                     & \cellcolor[HTML]{99EE99}6
                                     & \cellcolor[HTML]{99FF99}5
                                     & \cellcolor[HTML]{99DD99}7
                                     & \cellcolor[HTML]{999999}11
                                     & \cellcolor[HTML]{99DD99}7\\
\textbf{Total} (total software pieces) & \cellcolor[HTML]{99FF99}\textbf{7}
                                     & \cellcolor[HTML]{99E899}\textbf{9}
                                     & \cellcolor[HTML]{99FF99}\textbf{7}
                                     & \cellcolor[HTML]{99D299}\textbf{11}
                                     & \cellcolor[HTML]{999999}\textbf{16}
                                     & \cellcolor[HTML]{99DD99}\textbf{10}\\
\cellcolor[HTML]{EEEEEE}\textbf{Robustness}& \multicolumn{6}{l}{\cellcolor[HTML]{EEEEEE}}\\
Specific components -- \texttt{R$_1$} & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99CC99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}2
                                     & \cellcolor[HTML]{999999}2
                                     & \cellcolor[HTML]{99CC99}1\\
  Specific interactions -- \texttt{R$_2$} & \cellcolor[HTML]{99EB99}2
                                     & \cellcolor[HTML]{99D699}3
                                     & \cellcolor[HTML]{99FF99}1
                                     & \cellcolor[HTML]{99C299}4
                                     & \cellcolor[HTML]{999999}6
                                     & \cellcolor[HTML]{99D699}3\\
  \textbf{Total} (execution software pieces)& \cellcolor[HTML]{99F099}\textbf{2}
                                     & \cellcolor[HTML]{99D399}\textbf{4}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{99B699}\textbf{6}
                                     & \cellcolor[HTML]{999999}\textbf{8}
                                     & \cellcolor[HTML]{99D399}\textbf{4}\\
\cellcolor[HTML]{EEEEEE}\textbf{Extensibility}& \multicolumn{6}{l}{\cellcolor[HTML]{EEEEEE}}\\
  New engine type -- \texttt{E$_1$}  & \cellcolor[HTML]{99C599}3
                                     & \cellcolor[HTML]{99A899}4
                                     & \cellcolor[HTML]{99E299}2
                                     & \cellcolor[HTML]{99C599}3
                                     & \cellcolor[HTML]{999999}4.5
                                     & \cellcolor[HTML]{99FF99}1\\
New engine version -- \texttt{E$_2$} & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1\\
  New workflow -- \texttt{E$_3$} & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{99FF99}2
                                     & \cellcolor[HTML]{999999}3\\
New infrastructure -- \texttt{E$_4$} & \cellcolor[HTML]{99DD99}4
                                     & \cellcolor[HTML]{99DD99}4
                                     & \cellcolor[HTML]{99FF99}3
                                     & \cellcolor[HTML]{99DD99}4
                                     & \cellcolor[HTML]{999999}6
                                     & \cellcolor[HTML]{99DD99}4\\
  \textbf{Total} (difficulty to extend) & \cellcolor[HTML]{99D299}\textbf{10}
                                     & \cellcolor[HTML]{99D299}\textbf{10}
                                     & \cellcolor[HTML]{99FF99}\textbf{8}
                                     & \cellcolor[HTML]{99E899}\textbf{9}
                                     & \cellcolor[HTML]{999999}\textbf{12.5}
                                     & \cellcolor[HTML]{99E899}\textbf{9}\\
\cellcolor[HTML]{EEEEEE}\textbf{Scalability}& \multicolumn{6}{l}{\cellcolor[HTML]{EEEEEE}}\\
Multiple engine instances -- \texttt{S$_1$}& \cellcolor[HTML]{999999}2
                                     & \cellcolor[HTML]{99CC99}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99CC99}1
                                     & \cellcolor[HTML]{99CC99}1\\
Distributed engines -- \texttt{S$_2$}& \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1\\
Task scheduling -- \texttt{S$_3$}    & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0\\
\textbf{Total} (scalability issues)  & \cellcolor[HTML]{999999}\textbf{3}
                                     & \cellcolor[HTML]{99CC99}\textbf{2}
                                     & \cellcolor[HTML]{99CC99}\textbf{2}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{99CC99}\textbf{2}
                                     & \cellcolor[HTML]{99CC99}\textbf{2}\\
\cellcolor[HTML]{EEEEEE}\textbf{Specific features}& \multicolumn{6}{l}{\cellcolor[HTML]{EEEEEE}}\\
  Meta-workflow  -- \texttt{O$_1$}    & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{99FF99}0\\
  Fine-grained debugging -- \texttt{O$_2$}   & \cellcolor[HTML]{99FF99}0
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1
                                     & \cellcolor[HTML]{999999}1\\
  \textbf{Total} (missing features) & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{999999}\textbf{2}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{999999}\textbf{2}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}
                                     & \cellcolor[HTML]{99FF99}\textbf{1}\\
\end{tabular}

\caption{Architecture evaluation. Lower values (brighter colors) indicate better performance. Cell color is set as follows: (1) on each row, metric values are
  normalized between 0 (best value) and 1 (worst value):
  $m'=\frac{m-m_{\mathrm{min}}}{m_{\mathrm{max}}-m_{\mathrm{min}}}$ where
  $m$ is the metric value, $m_{\mathrm{min}}$ and $m_{\mathrm{max}}$
  are the minimum and maximum values for all architectures; (2) the RGB hexadecimal color code of the cell
  is \texttt{\#99XX99}, where X=\texttt{round}(F-6m') (\texttt{round} rounds a number to the nearest integer). }
\label{table:evaluation}
\end{table*}

\subsection{Evaluation metrics}

Science gateway \emph{integration complexity} is measured through the number of software components and their interactions. It breaks
down to the following 2 metrics:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Total components (\texttt{I$_1$}): total number of components in
  the architecture.
\item Total interactions (\texttt{I$_2$}): total number of
  interactions in the architecture.
\end{itemize}
One may wonder whether infrastructure should be counted in
\texttt{I$_1$}, since it is usually an external entity that is not developed nor maintained by the groups who
integrate workflow engines in science gateway. We support its inclusion here because integrating an
infrastructure into the system usually requires some technical
effort (e.g., account creation, software installation, APIs, etc.), therefore increasing complexity. The two metrics are summed to obtain a
global score that measures the
total number of software pieces affected by integration.

\emph{Robustness of workflow execution} measures the likelihood that the
workflow execution fails due to errors in the components or in the
interactions of the software architecture. Errors coming
from the infrastructure (e.g., unavailable data or terminated tasks)
or workflows (e.g., wrong user input or application errors) are not
covered, since they do not stem directly from the software
architecture. 
\closednote{vladimir}{I believe, proper error handling through the different levels of wf engine/gateway must be especially mentioned even if not especially covered. One of serious issues we faced while working with gateways at the AMC, in particular with gUse/WS-PGRADE, was that when it failed sometimes it was completely not clear what was the problem exactly. So we might consider discussing not only the number of software components but how explicit and clear interactions are between them; how transparently we can monitor the execution through all the levels and thus detect and prevent possible malfunction}{Tristan}{In our current evaluation framework, this is supposed to be covered by O2. I still agree with you that O2 could be further refined. See Github issue #27}
Robustness is measured here as a consequence of global
complexity, since complex architectures tend to be more prone to
failure. More precisely, robustness is determined by the number of
software components and interactions represented in
red in the architecture diagrams in
Figure~\ref{fig:architectures}, which are specific to workflow
execution. Two metrics are used:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Specific components  (\texttt{R$_1$}): number of specific components
involved in workflow executions. The Science
gateway component, for instance, is \emph{not} specific to workflow execution
since it is used to authenticate users, add new workflows, transfer
user data, etc., and
\item Specific interactions (\texttt{R$_2$}): number of specific
  interactions involved in workflow executions. Data control
  (interaction \texttt{c}), for instance, is \emph{not} specific to
  workflow execution since it is also used by the science gateway to
  transfer user data. 
\end{itemize}
\texttt{R$_1$} and \texttt{R$_2$} are
summed to obtain the global score for this criterion, which measures
the total number of software pieces that are specifically involved in
the workflow execution.

\emph{Extensibility} measures the difficulty to replace or add
elements to the architecture. It is determined by the number of interactions and
components that need to be modified when a new element is
added. Modification of a component is required when its code needs to
be updated or recompiled (science gateway or workflow service), or
when a new piece of software has to be installed (infrastructure
only). Modification of an interaction is deemed necessary when the
parameters involved in this interaction are modified.  Extensibility
breaks down to 4 metrics depending on the type of element (engine, workflow or infrastructure) that has to
be added or replaced:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item New engine (\texttt{E$_1$}): number of interactions or
  components that need to be modified to integrate a new type of
  workflow engine in the architecture. Workflow engines belong to
  different types when they cannot be invoked using the same
  implementation of the interactions.
 \item Engine version upgrade (\texttt{E$_2$}): number of interactions
  or components modified to integrate a new version of a workflow
  engine in the architecture, assuming that another version of the
  same engine type is already available. Different versions of a
  workflow engine can be invoked using the same interaction
  implementation(s). When this is not the case, the different versions
  are considered as different engine types. Components are not treated
  equally regarding engine version updates. Components whose only
  function is to host the workflow engine, i.e. workflow services and
  agents, are not counted in E$_2$ because updates in such components
  are assumed to be straightforward. On the contrary, modifications of
  components that have other functions than wrapping the engine,
  i.e. science gateway and infrastructure, are counted in E$_2$
  because updates in these components require more effort, e.g. new
  release of the science gateway, gaining administrative privileges on
  the infrastructure, etc. In practice, E$_2$=0 when the workflow
  engine is wrapped in a service or in an agent, and E$_2$=1 when it is
  wrapped in another component.  
\item New workflow (\texttt{E$_3$}): number of interactions required
  to integrate a new workflow in the architecture. Adding a new
  workflow is a very common operation that does not require modifying
  software components or interactions, assuming that the engine type
  and version required to execute this workflow are already
  available. In most cases, adding a new workflow requires only
  interactions \texttt{a$_1$} and \texttt{a$_2$}, but \texttt{g}
  is required as well when workflow conversion is used. 
\item New infrastructure (\texttt{E$_4$}): number of interactions or
  components affected for the integration of a new type of infrastructure in
  the architecture. Adding a new infrastructure typically aims at providing more
  computing or storage power, enabling access to specific types of resources
  (e.g., GPUs, clouds), or  enforcing execution policies (e.g.,
  constrain data to remain in a particular network domain).
%\item New science gateway: \todo{Not sure if this makes sense}.
\end{itemize}
The four metrics are summed to obtain a global score that measures the
difficulty to extend the architecture. Note that some extensions may
involve several metrics in practice. For instance, adding a new type
of engine may help integrating new infrastructures when interactions
\texttt{b} and \texttt{c$_2$} are already present for the new engine
and the infrastructure.

\emph{Scalability} corresponds to the ability of an architecture to
cope with high workloads through the following mechanisms: starting multiple engine instances, distributing engines and task scheduling. 
This is measured by assessing the potential
scalability difficulties due to (partial) absence of a given feature in the architecture.  
Features are evaluated using a 3-level metric: 0 means that
the feature is very easy to enable, 1 means that it can be implemented
but with some difficulty, and 2 means that the feature cannot
 be implemented in a realistic scenario. Three different features are identified:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Multiple engine instances (\texttt{S$_1$}): measures the
  potential for a single gateway instance to use more than one engine
  instance simultaneously. Workflow engines may require important
  amounts of resources when several workflows, or large workflows, are
  executed. At some point, it may be required for the science gateway
  to distribute the load among several engines. \texttt{S$_1$}=0 means
  \closednote{vladimir}{is this the right word here, 'to
    means'?}{Tristan}{This sentence had been truncated, thanks.} that
  adding a new engine instance is inherently supported in the
  architecture. In this case, elastic engines can be implemented
  through some kind of auto-scaling mechanism to control the number of
  engine instances in the architecture. \texttt{S$_1$}=1 means that
  multiple engine instances requires specific developments in the
  science gateway or workflow engine. \texttt{S$_1$}=2 means that new
  engine instances cannot be added.
\item Distributed engines (\texttt{S$_2$}): measures the potential of distributing
  the execution \emph{of a single workflow} among different engine
  instances. In our scope, this feature focuses on the capabilities of
  the architecture rather than these of the workflow
  engine. \texttt{S$_2$}=0 means that distributed engines are enabled by
  the architecture, and \texttt{S$_2$}=1 that they require specific
  developments in the workflow engine. 
\item Task scheduling (\texttt{S$_3$}): measures the complexity
  added by the architecture to task scheduling on
  the infrastructure. Task
  scheduling typically depends more on the implementation of specific
  algorithms in the science gateway, workflow engine and
  infrastructure than on the architecture used to integrate the
  workflow engine in the science gateway. Some architectures, however,
  complicate the task scheduling problem by introducing additional
  software layers or creating tasks with specific
  characteristics. \texttt{S$_3$}=0 means that the architecture does
  not add any additional complexity to the scheduling problem,
  and \texttt{S$_3$}=1 otherwise.
\end{itemize}
These three metrics are summed to obtain a global measure of the scalability
potential of the architecture.

\emph{Specific features} include:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Meta-workflow (\texttt{O$_1$}): measures the ability to describe
  meta-workflows from existing workflows. Meta-workflows offer an
  additional level of flexibility to build workflows from reusable
  components. \texttt{O$_1$}=0 means that meta-workflows are
  intrinsically enabled by the architecture (i.e. they can be
  implemented using the components and interactions already in place),
  amd \texttt{O$_1$}=1 means that they may be implemented with some
  development effort. \closednote{si}{why here we do not have the
    "possible to implement" situation?}{Tristan}{O$_1$=2 would never
    be reached. Besides, that would make it difficult to sum O1 and O2
    because they wouldn't measure the same type of thing. In the
    current situation, we are only talking about the presence/absence
    of a particular feature and this is why O1 and O2 can be added to
    obtain a total number of 'missing features'.}
  \closednote{vladimir}{I agree with Silvia, we should have the
    intermediate values as well, since meta-workflows might be
    straightforwardly supported by the gateway, or they are still
    possible to implement but only with some workarounds. Like it was
    with SHIWA: first 6 months coarsed-grained interoperability looked
    more like a workaround till it was better integrated and supported
    by WS-PGRADE itself}{Tristan}{Yes but in the situation you
    describe, evolutions came from the wf engine implementation rather
    than from the architecture to integrate it, which is our point
    here.}

\item Fine-grained debugging (\texttt{O$_2$}): availability of
  fine-grained debugging information about workflow tasks (white-box
  workflow).  Fine-grained information about workflow tasks is
  required to properly troubleshoot workflow executions. Its
  availability and accuracy usually depends on the number of software
  layers between the science gateway and the workflow
  engine~\cite{olabarriaga2014}. \texttt{O$_2$}=0 means that
  the information is directly accessible in the science gateway,
  \texttt{O$_2$}=1 means that it is obtained through one or more software
  interactions. \closednote{vladimir}{otherwise - does it mean the
    fine-grained info is available but hard to access or not available
    it all? the latter happens as well}{Tristan}{Precised the text.}
\end{itemize}
\texttt{O$_1$} and \texttt{O$_2$} are summed to obtain a total number of missing specific
features in the architecture.

\closednote{sil}{since we are dreaming, it seems to me that keeping provenance properly is also something to consider. the architecture can hide some details, and make it nearly impossible to keep provenance without hard coded hacks across the layers/components - i think}{Tristan}{Fixed the definition according to the discussion in PR #20.}

%"A Formal Approach to Support Interoperability in Scientific
%Meta-workflows" (reviewed for IWSG and JoGC) has a formal model to
%evaluate CGI and FGI.
%"four major approaches for workflow interoperability include a,b,c,d" (see Terstyanzky et al, "Enabli%ng scientific workflow sharing through CGI...", FGCS 2014) -> read this ref.
% http://www.oreilly.com/programming/free/files/software-architecture-patterns.pdf

The architectures described in Figure~\ref{fig:architectures} are
evaluated along these metrics in the remainder of this Section.

\subsection{Tight integration}

\paragraph{Integration complexity} This architecture does not require any
component in addition to the science gateway and infrastructure
(\texttt{I$_1$=2}) and it involves 5 interactions: \texttt{a$_1$},
\texttt{a$_2$}, \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{I$_2$=5}).

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), but interactions \texttt{b} and \texttt{c$_2$} are
(\texttt{R$_2$}=2).

\paragraph{Extensibility} Integrating a new type of workflow engine
requires modifying the science gateway as well as interactions
\texttt{b} and \texttt{c$_2$} (\texttt{E$_1$}=3). Updating a workflow
engine version requires modifications in the science gateway
(\texttt{E$_2$}=1).  Inserting a new workflow is done through
interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$}=2). Adding a new infrastructure generates updates in
interactions \texttt{a$_1$}, \texttt{b}, \texttt{c$_1$} and
\texttt{c$_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} Adding a new engine instance requires a new
instance of the complete science gateway
%, which is in general not possible
\closednote{sil}{why not? i agree that S=2, but i do not agree that it would be impossible. you say how below}{Tristan}{Updated the definition of S$_1$ to explain that multiple engine instances have to be available in the same sg instance.}
(\texttt{S$_1$}=2). 
Specific IT setups such as load-balancing between
web server instances might be used to create new gateway instances, but they would not allow a single gateway instance to use multiple engine instances, which is the point here.
Distributed engines are not available by default
(\texttt{S$_2$}=1). 
The architecture does not add any particular complexity to the task scheduling problem since the workflow engine
may implement any kind of scheduling policy (\texttt{S$_3$}=0).
\closednote{sil}{so same as the others = "easy"? i think this requires more explanation. maybe we should introduce a score for "dont know"?}{Tristan}{Fixed as discussed in PR #20}

\paragraph{Specific features} Supporting meta-workflows requires
specific developments in the workflow engine or science gateway to
allow workflows to execute other workflows (\texttt{O$_1$}=1). For
instance, specific Java objects may be implemented in the science
gateway framework to execute different types of workflows and link
them together.  \closednote{sil}{why not? if the engine is ws-pgrade
  it does. i really think that meta-workflows (and nested workflows)
  is a property exclusively of the engine. you can see that in fig 2.V
  interaction *6}{Tristan}{If the engine is WS-PGRADE as used in SHIWA
  CGI, then the architecture is nested workflows with service
  invocation. In a tight integration, meta-workflows would be
  expressed, e.g. as compositions of Liferay portlets that all
  implement different types of engines.}  The science gateway can
retrieve fine-grained debugging information from the workflow engine
directly (\texttt{O$_2$}=0).

\subsection{Service invocation}

\paragraph{Integration complexity} Service invocation requires a workflow service
in addition to the science gateway and infrastructure
(\texttt{I$_1$}=3). The architecture involves 6 interactions:
\texttt{a$_1$}, \texttt{a$_2$}, \texttt{b}, \texttt{c$_1$},
\texttt{c$_2$} and \texttt{d} (\texttt{I$_2$}=6).

\paragraph{Robustness} The workflow service is a component specific to
workflow execution (\texttt{R$_1$}=1). Workflow execution also
involves 3 specific interactions: \texttt{b},
\texttt{c$_2$} and \texttt{d} (\texttt{R$_2$}=3).

\paragraph{Extensibility} Adding a new type of workflow engine
requires implementing the corresponding workflow service,  modifying
interaction \texttt{d}, and  implementing interactions \texttt{b} and
\texttt{$c_2$} (\texttt{E$_1$}=4). New engine versions can be added by
updating the workflow service without modifying any interaction or
component (\texttt{E$_2$}=0). Updating the engine version in a
workflow service does not count in \texttt{E$_2$} since the only goal of this component
is to wrap the engine.  New workflows are added in the science gateway
or in the workflow engine through interactions \texttt{a$_1$} and
\texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new type of infrastructure
requires updates in interactions \texttt{a$_1$}, \texttt{b},
\texttt{$c_1$} and \texttt{$c_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} The service architecture in principle supports
multiple engine instances through multiple workflow services. Adding a
new engine instance, however, requires specific developments in the science gateway
(\texttt{S$_1$}=1). Distributing the execution of a single workflow in multiple
engines is usually not possible unless the workflow engine has
specific abilities (\texttt{S$_2$}=1).  \closednote{sil}{is this score for
  VIP? i think that the score depends on the implementation of the
  science gateway, which could adopt some elastic mechanism within. so
  we should not say s_2=1 for the architecture}{Tristan}{Removed VIP example. Updated score definition according to the discussion in @SilviaOlabarriaga's PR (\#20).} The architecture does not add any particular complexity to the task scheduling problem
since the workflow engine may implement any kind of scheduling
policy (\texttt{S$_3$}=0).  \closednote{sil}{see comment in section 3.1}{Tristan}{Fixed the same way}

\paragraph{Specific features} Meta-workflows require specific
developments in the workflow engine or science gateway to allow
workflows to execute other workflows.  (\texttt{O$_1$}=1).
\closednote{sil}{see comment in section 3.1}{Tristan}{same answer} The
science gateway needs to invoke interaction \texttt{d} to retrieve
dine-grained debugging information from the workflow service
(\texttt{O$_2$}=1).  \closednote{sil}{this is really not the case. the
  fact that the wf service is a separate component that hides
  information is a big problem for debugging at the science gateway
  level.  we had a paper about this:
  @inproceedings{olabarriaga2014scientific, Author = {Olabarriaga,
      Silvia and Pierantoni, Gabrielle and Taffoni, Giuliano and
      Sciacca, Eva and Jaghoori, Mahdi and Korkhov, Vladimir and
      Castelli, Giuliano and Vuerli, Claudio and Becciani, Ugo and
      Carley, Eoin and others}, Booktitle = {e-Science (e-Science),
      2014 IEEE 10th International Conference on}, Organization =
    {IEEE}, Pages = {298--305}, Title = {Scientific Workflow
      Management--For Whom?}, Year = {2014}} }{Tristan}{I agree. Fixed
  it and precised the definition of O2.}

\subsection{Task encapsulation}

\paragraph{Integration complexity} Task encapsulation requires only 2
components (\texttt{I$_1$}=2).  It involves 5 interactions:
\texttt{a$_1$}, \texttt{a$_2$}, \texttt{b}, \texttt{c} and
\texttt{e} (\texttt{I$_2$}=5).

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), and only interaction \texttt{e} is necessary
(\texttt{R$_2$}=1).
\closednote{vladimir}{there is a large piece of text in Scalability about reliability of the workflow engine task which on my opinion suits Robustness better}{Tristan}{I understand that that text may sound like robustness, but the robustness metrics are about the architecture, not the tasks, as explained in 3.1 (Errors coming from
the infrastructure (e.g., unavailable data or termi-
nated tasks) or workflows (e.g., wrong user input or
application errors) are not covered).}
\closednote{vladimir}{in addition, the claim of very good robustness of task encapsulation seems questionable to me as in this model you delegate the critical task of controlling workflow execution to, let's say, some 'random' node on the grid that you have no control of, not sure about its performance, connectivity and reliability and so on; however, considering the models and dependencies presented in fig.2 this evaluation of robustness metric may be correct}{Tristan}{You are right but the robustness issues that you are talking about are related to the infrastructure and not the architecture. This is really an issue related to task scheduling as the SG might decide to run the workflow tasks on reliable nodes of the infrastructure such as a dedicated cluster. So the issue that you mention is reflected in S3.}

\paragraph{Extensibility} Integrating a new type of workflow engine
requires developing interaction \texttt{e} and  installing the engine
on the infrastructure (\texttt{E$_1$}=2). Updating an engine version
in the architecture shares the same mechanism as version updates of
other tasks, which requires an update on the
infrastructure (\texttt{E$_2$}=1).  New workflows are integrated by
creating a new task in the science gateway through interactions
\texttt{a$_1$} and \texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new
infrastructure requires updating interactions \texttt{a$_1$},
\texttt{b} and \texttt{c} in the science gateway (\texttt{E$_4$}=3).

%\todo{Integrating a workflow engine as a sub-task
%only requires to implement interaction \texttt{e} (\texttt{I$_2$}=1)
%and to install the workflow engine on the infrastructure
%(\texttt{I$_3$}=1). A new task type is created in the science gateway
%only when a new workflow has to be integrated (see M$_4$), but this is
%not required to integrate the engine itself (\texttt{I$_1$}=0).}

\paragraph{Scalability}
New engine instances are spawned and executed on the infrastructure as
any other task upon user submission (\texttt{S$_1$}=0). This is a
major interest of task encapsulation. Distributed engines are not
supported by default (\texttt{S$_2$}=1). Task scheduling is slightly
more complex than in the other approaches due to the special role of
the task that executes the workflow engine (\texttt{S$_3$}=1). 
Indeed,
the reliability of this task is critical since all the sub-tasks in
the workflow depend on it and, depending on the recovery capabilities
of the workflow engine, may need to be resubmitted if the workflow
task fails. The workflow task is also longer than all its sub-tasks,
which increases its chances of failure. \closednote{vladimir}{isn't this about robustness?}{Tristan}{Nope, this is task scheduling, therefore scalability in our evaluation framework. See answer to your previous note.} In addition, task parameters,
for instance estimated walltime, are more difficult to estimate for
the workflow task than for the sub-tasks because workflows are by
definition more complex than their sub-tasks: errors on sub-task
parameter estimations accumulate in the workflow, and additional
control constructs such as tests and loops may further increase the
uncertainty. Such parameter estimation errors may generate issues such
as selection of wrong batch queues on clusters or task termination due
to exceeded quotas. Finally, the interdependencies between
the workflow task and its sub-tasks may create deadlocks when there is
contention. For instance, if only 1 computing resource is available
for the science gateway and if the workflow task is running on it and
submits sub-tasks, then the sub-tasks could only execute when the
resource is available, which will never happen because the workflow
task will not complete until the sub-tasks complete. This
configuration can be generalized to an infrastructure with $n$
resources where $n$ workflows are submitted. In practice, however, the
number of submitted workflows usually remains lower than the number of
computing resources available on this infrastructure, which makes such
deadlocks unlikely to happen.

\paragraph{Specific features} Task encapsulation explicitly enables
meta-workflows (\texttt{O$_1$}=0). Indeed, workflow engines may invoke
other engines using the exact same interaction used to submit regular
sub-tasks (interaction \texttt{e}), without any additional
development. \closednote{sil}{see comment in section
  3.1}{Tristan}{Fixed, actually updated the score of this architecture
  now that the definition is more accurate.} Fine-grained debugging
information is obtained through interaction \texttt{b}
(\texttt{O$_2$}=1). \closednote{sil}{mm, i think the case here is the
  same as for service. or actually i would give this one a better
  score, because there are less components to cross}{Tristan}{Fixed the text to better match the updated definition of O2}

\subsection{Pool model}

\paragraph{Integration complexity} The pool model requires a workflow pool and an
agent in addition to the science gateway and infrastructure
(\texttt{I$_1$=4}). It involves 7 interactions: \texttt{a$_1$},
\texttt{a$_2$}, \texttt{b}, \texttt{c$_1$}, \texttt{c$_2$}, \texttt{d}
and \texttt{f} (\texttt{I$_2$}=7).

\paragraph{Robustness} The workflow pool and agent are specific to
workflow execution (\texttt{R$_1$=2}). Interactions \texttt{b},
\texttt{c$_2$}, \texttt{d} and \texttt{f} also are (\texttt{R$_2$=4}).

\paragraph{Extensibility} Adding a new engine type requires wrapping
the engine into the agent and updating interactions \texttt{b} and
\texttt{c$_2$} (\texttt{E$_1$=3}). Updating the version of an engine
is transparent (\texttt{E$_2$=0}) since it only requires updating the
agent, which is a component dedicated to the engine. Integrating a new workflow is
done through interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$=2}). Integrating a new infrastructure requires updates
in interactions \texttt{a$_1$}, \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{E$_4$=4}).

\paragraph{Scalability} By design, new engine instances only require
new agents, which can be easily automated (\texttt{S$_1$=0}). For
instance, auto-scaling rules can be implemented to start new agents
when the workload in the science gateway exceeds a certain
threshold~\cite{lorido2012auto}. Distributed engines are not available
by default (\texttt{S$_2$=1}) and the architecture does not add any
complexity to the task scheduling problem (\texttt{S$_3$=0}).

\paragraph{Specific features} Meta-workflows require specific
developments in the workflow engine to enable workflow submission
(\texttt{O$_1$=1}).  \closednote{sil}{see comment in section 3.1}{Tristan}{Same answer.} Debugging information is accessed through interactions \texttt{d} and \texttt{f} (\texttt{O$_2$=1}).
\closednote{sil}{i think the case here is the same as for service. or
  actually i would give this one a worse score, because there is one
  additional component - the pool - to cross}{Tristan}{Fixed using the updated definition of O2}

\subsection{Nested workflows with service invocation}

\paragraph{Integration complexity} Setting up a nested workflow
architecture with service invocation requires a science gateway, 2
workflow services and 2 infrastructures (\texttt{I$_1$=5}). The
architecture involves 11 interactions (\texttt{I$_2$=11}):
\texttt{a$_1$} (twice), \texttt{a$_2$}, \texttt{b} (twice), \texttt{c$_1$} (twice),
\texttt{c$_2$} (twice) and
\texttt{d} (twice).
\closednote{vladimir}{and if we have a meta-workflow with 3 and more sub-workflows the score will be even higher, is it a fair calculation for an architecture? we did not increase the scores for possible meta-workflows supported by wf engines in other cases}{Tristan}{Good point! You are totally right for meta-workflows, I added a sentence in the paragraph below. However, nested workflows are studied here as an architecture to integrate a workflow engine (and only one), not to describe meta-workflows.}

\paragraph{Robustness} The two workflow services are specific to
workflow execution (\texttt{R$_1$=2}). Interactions \texttt{b}
(twice), \texttt{c$_2$} (twice) and \texttt{d} (twice) also
are necessary (\texttt{R$_2$=6}).

\paragraph{Extensibility} Adding a new type of \emph{parent} engine
requires implementing the corresponding service,  implementing
interactions \texttt{b} and \texttt{c$_2$} in the parent engine, and
implementing interaction \texttt{d} in the science gateway and in the
parent service (\texttt{E$_1$=5}). Adding a new type of \emph{child}
engine only requires implementing the corresponding service, 
developing interactions \texttt{b} and \texttt{c$_2$} in the child
engine, and implementing interaction \texttt{d} in the parent service
(\texttt{E$_1$=4}). We use \texttt{E$_1$=4.5} in
Table~\ref{table:evaluation} to reflect both conditions. Adding a new
version in the parent or child engine only requires modifying this
engine (\texttt{E$_2$=0}). Adding a new workflow is done through
interaction \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$=2}). Adding a new infrastructure requires 
re-implementing interactions \texttt{a$_1$}, \texttt{b} and
\texttt{c$_2$} twice, and interaction \texttt{c$_1$} once so that it
can be supported by both workflow engines (\texttt{E$_4$=6}).

\paragraph{Scalability} As in the service architecture, adding a new
engine instance requires specific developments in the science gateway
(instance of a parent engine), or in the parent engine (instance of a
child engine) (\texttt{S$_1$=1}). Similarly, elastic engines are
difficult to achieve. Distributed engines can be implemented through meta-workflows (\texttt{S$_2$=0}). Task scheduling
is more complex than in other architectures though, due to the fact
that workflow execution is split in different engines (\texttt{S$_3$=1}).

\paragraph{Specific features}
Meta-workflows can be implemented, which is one of the main interest
of this architecture (\texttt{O$_1$=0}). Note, however, that the
complexity of the architecture increases with the number of engine
types involved in meta-workflows: for instance, a meta-workflow with
child workflows executed by 2 different engine types would require an
additional workflow service and the corresponding interactions.
Debugging information is accessed through interaction \texttt{d} (invoked twice) (\texttt{O$_2$=1}).

\subsection{Workflow conversion with service invocation}

\paragraph{Integration complexity} Workflow conversion with service invocation
requires the same components as for service invocation
(\texttt{I$_1$}=3), and an additional \texttt{g} interaction (\texttt{I$_1$}=7).

\paragraph{Robustness} Since workflow conversion is not involved in
the execution (it is an offline process), the scores are the same as for the
service architecture (\texttt{R$_1$=1}, \texttt{R$_2$=3}).

\paragraph{Extensibility} Since adding a new type of workflow engine
aims at supporting more workflows, we consider that it only requires
re-implementing interaction \texttt{g} in this architecture  (\texttt{E$_1$=1}). Note,
however, that implementing interaction \texttt{g} can require
substantial work depending on the complexity of the workflow language used by
the new engine. Similarly, adding a
new engine version only requires modifying interaction \texttt{g}
(\texttt{E$_2$=1}).  Adding a new workflow is done through
interactions \texttt{a$_1$}, \texttt{a$_2$} and \texttt{g} (\texttt{E$_3$=3}). As in the
service architecture, interfacing with a new infrastructure requires
modifications in interactions \texttt{a$_1$}, \texttt{b}, \texttt{c$_1$}
and \texttt{c$_2$} (\texttt{E$_4$=4}).

\paragraph{Scalability}  Since workflow conversion is not involved in
the execution (it is an offline process), the score is the same as for the
service architecture (\texttt{S$_1$=1},
\texttt{S$_2$=1}, \texttt{S$_3$=0}).

\paragraph{Specific features} Meta-workflows are available after
conversion, by connecting workflows in the language used in the
science gateway (\texttt{O$_1$=0}).  \closednote{sil}{i dont
  understand this about meta-workflows. translation does not imply
  composition, right?  if the science gateway is allowed to implement
  workflow composition, then this is probably the case in other
  architectures as well?}{Tristan}{Well at this level, connecting two
  workflows expressed in the same language is straightforward,
  although I agree that it might be tricky in some cases.} Debugging information is accessible
as in the service invocation architecture (\texttt{O$_2$=1}).
\closednote{sil}{see notes 3.2}{Tristan}{Fixed the same way}

\section{Discussion}

\subsection{Comparison between architectures}

Tight integration and task encapsulation are the simplest
architectures to integrate, followed by service integration, workflow conversion (with
service invocation) and pool. Nested workflows (with service
invocation) require more integration than the other
architectures. Robustness roughly leads to the same ordering of
architectures, with tight integration and task encapsulation in the top
group, service integration and workflow conversion (with service
invocation) close behind, pool in a third group, and nested workflows
(with service invocation) at the end. This ordering is consistent
across metrics; it reflects the global complexity of the
architectures.

Regarding extensibility, most architectures are overall comparable,
except nested workflows (with service invocation) which are
significantly behind. This is explained by the complexity of the
nested workflow architecture, with 2 infrastructures and 2 workflow
services. Task encapsulation and workflow conversion (with service
invocation) perform slightly better when integrating new engines
(\texttt{E$_1$}), which makes them useful architectures for science
gateways that do not focus on a particular engine.  However, task encapsulation, workflow conversion (with
service invocation) and tight integration perform worse than the
others when adding engine versions (\texttt{E$_2$}), due to the need
to update infrastructure (task encapsulation), science gateway (tight
integration), or workflow converter (workflow conversion). Workflow conversion
performs worse than the others when integrating new workflows
(\texttt{E$_3$}) because of the language conversion step. All
architectures except nested workflows are equivalent when integrating
new infrastructures (\texttt{E$_4$}).

The pool architecture is overall the most scalable, which is not
surprising since it is designed precisely for scalability. Tight
integration is the least scalable and all the other architectures
perform the same overall. The global scalability score, however,
should not conceal the unique characteristics of architectures regarding
this criterion. Nested workflows are the only architecture that can
easily accommodate distributed workflow execution, which can be
critical in some cases. At the same time, the scheduling constraints
created by task encapsulation and nested workflows may become problematic  depending on the type of targeted
infrastructure. Non-reliable infrastructures,
for example, could hardly cope with workflow engines being wrapped in
computing tasks as done in task encapsulation. The availability of
multiple engine instances could also become a critical feature for
science gateways with important workloads, which would favor task
encapsulation and pool. \note{Naj}{I feel this paragraph is a bit
  vague, and doesn't distill and contextualize the thorough work
  described before.}

Differences in other specific features should not be neglected. Task encapsulation, nested
workflows (with service invocation) and workflow conversion (with service
invocation) are the only architectures that intrinsically support meta-workflows. \closednote{vladimir}{As mentioned earlier by Silvia, this might not be exactly the case, depending on the point of view}{Tristan}{The example given by Silvia (WS-PGRADE) actually comes down to the nested workflow architecture. I added 'intrinsically' to the sentence to better reflect that.} Besides,
task encapsulation and nested workflows may hamper
fine-grained debugging. 
The availability of fine-grained debugging information may also be
critical to efficient user support.

Table~\ref{table:overall} provides an overall comparison between
architectures, based on the metrics in Table~\ref{table:evaluation}.
Note that this analysis is only meant for illustration purposes, since
it assumes that all criteria have equal weights while real systems
would favor some of them. Overall, task encapsulation performs a bit
better than the other architectures and nested workflows stand a bit
behind for the reasons mentioned previously.
\begin{table*}
\footnotesize
\centering
\begin{tabular}{rcccccc}
                                    & \textbf{Tight}
                                    & \textbf{Service}
                                    & \textbf{Task}
                                    & \textbf{Pool}
                                    & \textbf{Nested}
                                    & \textbf{Conversion} \\
  Integration (global score)    &
                                    \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99E899}{0.22}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99D299}{0.44}
                                    & \cellcolor[HTML]{999999}{1.00}
                                      & \cellcolor[HTML]{99DD99}{0.33}\\
Robustness (global score) &
                                \cellcolor[HTML]{99F099}{0.14}
                                    & \cellcolor[HTML]{99D399}{0.43}
                                    & \cellcolor[HTML]{99FF99}{0.00}
                                    & \cellcolor[HTML]{99B699}{0.71}
                                    & \cellcolor[HTML]{999999}{1.00}
                                    & \cellcolor[HTML]{99D399}{0.43}\\
  Extensibility (global score)  & \cellcolor[HTML]{99D299}{0.44}
                                     & \cellcolor[HTML]{99D299}{0.44}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99E899}{0.22}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99E899}{0.22}\\
Scalability (global score)  & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99CC99}{0.50}
                                     & \cellcolor[HTML]{99CC99}{0.50}\\
Specific features (global score) & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{999999}{1.00}
                                     & \cellcolor[HTML]{99FF99}{0.00}
                                     & \cellcolor[HTML]{99FF99}{0.00}\\
                                    & \cellcolor[HTML]{99DA99}\textbf{1.6}
                                    & \cellcolor[HTML]{99B899}\textbf{2.6}
                                    & \cellcolor[HTML]{99FF99}\textbf{0.5}
                                    & \cellcolor[HTML]{99BF99}\textbf{2.4}
                                    & \cellcolor[HTML]{999999}\textbf{3.5}
                                    & \cellcolor[HTML]{99DE99}\textbf{1.5}\\
\end{tabular}
\caption{Overall evaluation. Brighter colors and lower scores indicate better performance. Scores
  are obtained by summing the normalized global scores ($m'$ values) of
  each criterion in Table~\ref{table:evaluation} and the colors are obtained as in Table~\ref{table:evaluation}. }
\label{table:overall}
\end{table*}

\subsection{Limitations}

A few limitations should be considered when using the results of our
evaluation. First, our evaluation methods aimed to provide comparative
metrics, however the high-level of abstraction used to derive these
metrics hinders the complexity of real systems to some
extent. In particular, the
presented architectures are abstract patterns that may be mixed
together in actual systems. The distinction between tight integration and
service invocation, for instance, may not always be that clear in
practice. Service invocation may also be combined with task
encapsulation in some cases. However, the criteria discussed in this
work would still apply to broadly compare and categorize such cases of
hybrid architectures.

Moreover, all the interactions involved in the architectures were
treated equally, whereas some of them are obviously more complex than
others. For example, interaction \texttt{g} (workflow language
conversion) is clearly more complex than interaction \texttt{d}
(service invocation), and may challenge reliability more strongly. However, without entering into the details of a particular
system, quantification of the robustness or the amount of development
associated with each interaction and software component will hardly be
precise.

The particular case of interaction \texttt{g} (workflow language
conversion) requires particular attention when implementing a real
system since this interaction may not be easily generalizable to any
workflow language. For instance, converting FSL, PSOM or Nipype
pipelines to any other workflow language is problematic because these
engines rely on general-purpose programming languages such as Bash,
Octave and Python, which are much richer than scientific workflow
languages.

The abstract nested workflows and workflow conversion patterns were
instantiated with service invocation so that they can be analyzed in
the same framework as the other architectures. Other types of
instantiation, for instance nested workflows with task encapsulation,
could also be envisaged. We chose to limit ourselves to instantiations
with service invocation because the resulting architectures are
implemented in real systems, and because service invocation is largely
used. Nevertheless, it could be interesting to explore other types of
instantiations, in particular nested workflows with task
encapsulation.\answerednote{Naj}{I think in the discussion, you may want to close the circle on a statement in introduction, which referred to your 7-year experience with CBRAIN and VIP}{Tristan}{Good idea, will think about it.}

The particular technical or historical context of a science gateway
project may obviously influence the choice of an architecture to
integrate workflow engines. For instance, some workflow engines may be
already available as web services\note{Naj}{This is a good place to offer some examples}, which would tend to favor service
invocation, and other science gateways may have strongly tested task
and data control features (interactions \texttt{b} and \texttt{c}),
which would favor task encapsulation. Similarly, adding a new type of engine
may facilitate integration of a new infrastructure when interactions
\texttt{b} and \texttt{c$_2$} are already available. The migration
cost between architectures has been ignored as well.

Finally, workflow engines are only one of the many aspects involved in
the design of a science gateway. Other properties definitely influence
the design of a software architecture, for instance collaborative
features, data visualization, search, authentication, accounting and
so on.

\section{Related work}
\label{sec:related}

The early work~\cite{yuTaxonomy} proposes a generic architecture for
grid workflow systems that is based on the workflow reference model
defined by the Workflow Management
Coalition\footnote{\url{http://www.wfmc.org}}.  Deelman et
al. \cite{deelman2009workflows} have further characterized the
features of workflow systems and proposed a general architecture??
These two works, however, do not mention science gateways.

Shahand et al.~\cite{shahand:2015ab} presents the Science Gateway
Canvas, which is a business reference model where the most relevant
functional components are organized into functional groups. Although
workflow management is mentioned as a possible component to coordinate
distributed computations in science gateways, the paper does not
comment on architecture to achieve this.

Olabarriaga et al.~\cite{olabarriaga2014} presents a user-centered
view of the ecosystem of science gateways with focus on workflow
management. Their proposed layered system architecture includes
gateway, workflow engine and distributed infrastructure components,
similarly to the "service invocation" pattern. However other patterns
are not identified.
\note{vladimir}{@silvia, does it make sense to mention workflow classification efforts based on workflow motifs and so on?}

\section{Conclusion}

% Summary of contributions
We have systematically reviewed architectures used to integrate workflow engines in
science gateways. These architectures are described in a system-independent framework
suitable for comparison, illustrated on real systems, and evaluated using novel quantitative metrics that allow simple comparison across architectures. We have discussed the
pros and cons of all the presented architectures based on these metrics.

% Impact of the work
To the best of our knowledge, our work is the first to systematically
review and compare software architectures used to integrate workflow
engines into science gateways. So far, the literature on science gateways and
workflow engines has focused on the description of particular systems,
or on the presentation of a particular architecture.  Instead, our
analysis abstracts and evaluates architectural patterns independently
from any particular system. It provides general insight to integrate
science gateways and workflow engines, which we hope will be useful
for software architects.

% Literature on SG architecture Very few works go beyond the
% description of architectures of particular systems. For instance, the work
%in~\cite{Gannon2005} discusses service-oriented architectures for
% science gateways with workflow engines.  The work in
% http://link.springer.com/chapter/10.1007/11596141_3 discusses
% service-oriented architectures for science gateways, with a workflow
% engine. The work in
% http://docs.lib.purdue.edu/cgi/viewcontent.cgi?article=1414&context=nanopub
% highlights 4 TeraGrid science gateways (GridChem, LEAD, nanoHUB,
% CABIG). Architectures are described but there is no real focus on
% workflow engines. GridChem doesn't include any workflow engine,
% which btw make it relevant for integration via sub-tasking. LEAD has
% a workflow engine.

\section{Acknowledgments}

This work has been made possible with the support of the Irving Ludmer
Family Foundation and the Ludmer Centre for Neuroinformatics and
Mental Health. The integration between PSOM and CBRAIN was supported
by a Brain Canada Platform Support Grant, as well as the Canadian
Consortium on Neurodegeneration in Aging (CCNA), through a grant from
the Canadian Institute of Health Research and funding from several
partners.  This work is in the scope of the LABEX PRIMES (ANR-11-
LABX-0063) of Universit\'e de Lyon, within the program
``Investissements d’Avenir'' (ANR-11-IDEX-0007) operated by the French
National Research Agency (ANR). This work also falls into the scope of
the scientific topics of the French National Grid Institute (IdG). The
VIP team thanks the site administrators of the European Grid
Initiative and the GGUS support for their help related to the VIP
platform. The CBRAIN team is grateful for the computing cycles,
storage, and support obtained from Compute Canada
(\url{https://computecanada.ca}) and platform development program from
CANARIE (\url{http://www.canarie.ca/}). We also acknowledge the Dutch
national e-Infrastructure with the support of SURF Cooperative, the
Dutch national program COMMIT/ and the High Performance Computing and
Networking (HPCN) Fund of the University of Amsterdam for their
support to the science gateway actitivies at the AMC.  We are also
grateful to the financial support provided by FP7 E-INFRASTRUCTURE
program for financial support to SCI-BUS, SHIWA and ER-flow
projects.

\section*{References}

\enlargethispage{5mm}
\bibliographystyle{plain}
\bibliography{biblio}

\end{document}
\endinput